---
title: "p8105_hw3_jl6647"
author: "Jiatong LI"
date: "2023-10-05"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
library(dplyr)
```
```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom")) 
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## HW3 Problem 1
```{r}
data("instacart")
```
The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.

##### Some description:
  This dataset `instacart` has `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row representing a single product from an order. Variables all related to orders or products. For example, variables like `product_name`, `aisle`, `department` are related to the information of products. 
  
  Key variables include `r names(instacart)`.
  
##### Questions:

*1.How many aisles are there, and which aisles are the most items ordered from?
```{r}
aisles_num = 
  instacart |> 
  group_by(aisle_id) |> 
  summarize(n_products = n()) 
```
```{r}
most_iterms = 
  instacart |> 
  group_by(aisle) |> 
  summarize(n_products = n()) |> 
  mutate(rank_aisle = min_rank(desc(n_products))) |> 
  arrange(rank_aisle) |> 
  filter(rank_aisle == 1) |> 
  pull(aisle)
```

There're `r nrow(aisles_num)` aisles. `r most_iterms` is the aisles the most iterms ordered from.

*2.Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

  
## HW3 Problem 2: BRFSS
```{r}
data("brfss_smart2010")
```

### Data cleaning:

format the data to use appropriate variable names;

focus on the “Overall Health” topic

include only responses from “Excellent” to “Poor”

organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r}
brfss_df = 
  brfss_smart2010 |>
  janitor::clean_names() |> 
  filter(
    topic == "Overall Health",
    response %in% c("Excellent","Very Good","Good","Fair","Poor")
  ) |>
  mutate(
    response = factor(
      response,
      levels = c("Poor","Fair","Good","Very Good","Excellent"),
      ordered = TRUE
    )
  ) |> 
  separate(
    col = locationdesc,
    into = c("state","county"),
    sep = "-",
    remove = TRUE
  ) |> 
  select(year,state,county,class,topic,question,response,sample_size,data_value,geo_location)
  view(brfss_df)
```
##### Answer questions

*1.In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
state_2002 = 
  brfss_df |> 
  filter(
    year == 2002,
    response == "Excellent") |> 
  group_by(year,state) |> 
  summarize(n_county = n()) |> 
  filter(n_county >=7) |> 
  pull(state)

state_2010 = 
  brfss_df |> 
  filter(
    year == 2010,
    response == "Excellent") |> 
  group_by(year,state) |> 
  summarize(n_county = n()) |> 
  filter(n_county >=7) |> 
  pull(state)

```
In 2002, the states observed at 7 or more locations are `r state_2002`
In 2010, the states observed at 7 or more locations are `r state_2010`

*2.Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}
excellent_df =
  brfss_df |> 
  filter(response == "Excellent") |> 
  group_by(year,state) |> 
  summarize(avg_value = mean(data_value, na.rm = TRUE))
excellent_df
```
```{r}
excellent_plot =
  ggplot(
    excellent_df,
    aes(x = year, y = avg_value, color = state, group = state)) +
    geom_line() +
    theme_minimal() +
    labs(title = "Spaghetti plot: Average value of excellent responses over time within a state",
         x = "Year",
         y = "Average value of excellent responses")
excellent_plot
```
```{r}
min_df = 
  excellent_df |> 
  filter(year == 2005)
view(min_df)
```

From the plot, it is clear to see that the plot of the state `WV` is significantly different from other states, with low values in the year of 2005 and 2009.

*3.Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.





  


