---
title: "p8105_hw3_jl6647"
author: "Jiatong LI"
date: "2023-10-05"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
```
```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom")) 
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## HW3 Problem 1
```{r}
data("instacart")
```
The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.

##### Some description:
  This dataset `instacart` has `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row representing a single product from an order. Variables all related to orders or products. For example, variables like `product_name`, `aisle`, `department` are related to the information of products. 
  
  Key variables include `r names(instacart)`.
  
##### Questions:

*1.How many aisles are there, and which aisles are the most items ordered from?
```{r}
aisles_num = 
  instacart |> 
  group_by(aisle_id) |> 
  summarize(n_products = n()) 
```
```{r}
most_iterms = 
  instacart |> 
  group_by(aisle) |> 
  summarize(n_products = n()) |> 
  mutate(rank_aisle = min_rank(desc(n_products))) |> 
  arrange(rank_aisle) |> 
  filter(rank_aisle == 1) |> 
  pull(aisle)
```

There're `r nrow(aisles_num)` aisles. `r most_iterms` is the aisles the most iterms ordered from.
  
## HW3 Problem 2: BRFSS
```{r}
data("brfss_smart2010")
```

### Data cleaning:

format the data to use appropriate variable names;

focus on the “Overall Health” topic

include only responses from “Excellent” to “Poor”

organize responses as a factor taking levels ordered from “Poor” to “Excellent”


  


